---
title: "Manipulating Strings"
author: "Sarah Wigodsky"
date: "September 11, 2017"
output: html_document
---

####Input Names
```{r input_names, eval=TRUE}
library("stringr")
name <- c("Moe Szyslak", "Burns, C. Montogomery", "Rev. Timothy Lovejoy", "Ned Flanders", "Simpson, Homer", "Dr. Julius Hibbert")
```

####Extract Last Names - Question 3A
```{r extract_last_names, eval=TRUE}
last_name <-unlist(str_extract_all(name, "[[:alpha:]]{2,},|[^[\\.]][[:blank:]][[:alpha:]]{3,}"))
last_name <-unlist(str_extract_all(last_name, "[[:alpha:]]{2,}"))
last_name
```

####Extract First Names
```{r extract_first_names, eval=TRUE}
first_name <-unlist(str_extract_all(name, ", [[:alpha:]]{2,}|[[:upper:]]{1}. [[:alpha:]]{2,}|[[:alpha:]]{3,}[[:blank:]]"))
first_name <-unlist(str_extract_all(first_name, "[[:alpha:]]{2,}|[[:upper:]]{1}. [[:alpha:]]{2,}"))
first_name
```
####Write Names in Data Frame
``` {r write_first_name_last_name, eval=TRUE}
name_table <- data.frame (first_name, last_name)
name_table
```

#### Logical Vector to indicate title Question 3B
```{r determine_presence_of_title, eval=TRUE}
title <- str_detect(name, "[[:upper:]]{1}[[:alpha:]]{1,}\\.")
name_table <- cbind(name_table, title)
name_table
```

####Logical Vector to indicate a second name Question 3C
```{r determine_presence_of_second_name, eval=TRUE}
two_names <- str_detect(first_name, "[[:upper:]]{1}\\. [[:alpha:]]{1,}")
name_table <- cbind(name_table, two_names)
name_table
```

####Question 4
Describe the types of strings that confom to the following regular expressons and construct an example that is matched by the regular expression.

a.  [0-9]+\\$

    a digit between 0 and 9 that is repeated one or more times followed by $
    
    example:  45$

```{r, eval=TRUE}
a <- "45$"
str_extract_all(a, "[0-9]+\\$")
```


b.  \\b[a-z]{1,4}\\b

    a string that begins with a lower case letter that appears at least once but no more than 4 times followed by where the number is the end of the string
    
    example: yes

```{r, eval=TRUE}
b <- "yes"
str_extract_all(b, "\\b[a-z]{1,4}\\b")
```
    
c.  .*?\\.txt$

    a single character matched zero or more times (the preceeding item is optional and will be matched at most once) followed by a period and followed by txt
    
    example: file.txt

```{r, eval=TRUE}
c <- "file.txt"
str_extract_all(c, ".*?\\.txt$")
```


    
d.  \\d{2}/\\d{2}/\\d{4}

    a two digit number followed by / followed by a 2 digit number followed by / followed by a 4 digit number ... a date
    
    example:  11/12/1985

```{r, eval=TRUE}
d <- "11/12/1985"
str_extract_all(d, "\\d{2}/\\d{2}/\\d{4}")
```

    
e.  <(.+?)>.+?</\\1>

    a string that starts with < which is followed by one or more characters and then followed by > which is followed by at least 1 character which is followed by the repeat of the first set of characters in sideways brackets preceeded by /
    
    example:   <h1>heading</h1>
    
```{r, eval=TRUE}
e <- "<h1>heading</h1>"
str_extract_all(e, "<(.+?)>.+?</\\1>")
```

    