---
title: "DATA607HW7"
author: "Sarah Wigodsky"
date: "October 14, 2017"
output:
  html_document: default
  word_document: default
---
##Introduction
In this assignment, I will be reading in an HTML file, XML file and JSON file that I wrote that lists 3 of my favorite books.  I will then use R to make a meaningful table with the data stored in each format.

####Importing Libraries
```{r library, eval=TRUE}
suppressWarnings(suppressMessages(library(XML)))
suppressWarnings(suppressMessages(library(stringr)))
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(RCurl)))
suppressWarnings(suppressMessages(library(RJSONIO)))
```


####Reading in HTML data
```{r inputHTMLdata, eval=TRUE}
htmllink <- getURL("https://raw.githubusercontent.com/swigodsky/DATA-607/master/sarahfavbooks.html")
bookhtml <- htmlParse(file=htmllink)
```

####Creating a Data Table from the HTML file
```{r datatablefromhtml, eval=TRUE}
getth = function(){
  th_container = character()
  list(th = function(node, ...){
    th_container <<- c(th_container, xmlValue(node))
  }, returnth = function() th_container)
}

h1 <- getth()
htmlTreeParse(file = htmllink, handlers = h1)

row <- rep(NA, 6)
bookhtml <- data.frame(stringsAsFactors = FALSE)
bookhtml <- rbind(bookhtml, row)
row <- NULL

for (i in 7:length(h1$returnth())){
    row <- c(row, h1$returnth()[i])
    if (i%%6==0) {
      bookhtml <- rbind(bookhtml, row)
      row <- NULL
     }
}
rownames(bookhtml) <- NULL
for (i in 1:6){
    row <- c(row, h1$returnth()[i])
}
colnames(bookhtml) <- row
bookhtml.df <- bookhtml[-1,]
bookhtml.df
```

####Loading XML file
```{r xmlfile, eval=TRUE}
xmllink <- getURL("https://raw.githubusercontent.com/swigodsky/DATA-607/master/sarahfavbooks2.xml")
bookxml <- xmlParse(file = xmllink)
```

####Creating a Data Table from the XML file
```{r xmltodataframe, eval=TRUE}
bookxml.df <- xmlToDataFrame(xmlRoot(bookxml))

authors <- xpathSApply(bookxml, "//author", xmlAttrs)
bookxml.df$author <- c(authors[1,1], authors[1,2], authors[1,3])
colnames(bookxml.df)[2] <- "author1"
bookxml.df <- cbind(bookxml.df, c(authors[2,1], authors[2,2], authors[2,3]))
colnames(bookxml.df)[6] <-"author2" 

bookxml.df <- bookxml.df %>%
  select(title, author1, author2, publisher, ISBN, award)
bookxml.df
```

####Loading JSON file
```{r jsonfile, eval=TRUE}
bookjson <- fromJSON("https://raw.githubusercontent.com/swigodsky/DATA-607/master/sarahfavbooks.json", nullValue = NA, simplify=FALSE)
```

####Creating a Data Table from the JSON file
```{r jsontodataframe, eval = TRUE}
bookjson.df <- do.call("rbind", lapply(bookjson, data.frame, stringsAsFactors=FALSE))

colnames(bookjson.df)[2] <- "author1"
colnames(bookjson.df)[3] <- "author2"
bookjson.df[is.na(bookjson.df)] <- "NULL"
bookjson.df
```