---
title: "DATA_607_Project3- Data Science Skills"
authors: "Jill Anderson, Alvaro Bueno Castillo, Nathan Cooper, Silverio Vasquez, Sarah Wigodsky"
date: "October 16, 2017"
output: html_document
---

##What are the most valued data science skills?
To answer this question, we mined the job board glassdoor.com to identify which skills were requested the most frequently.

####Importing Libraries
```{r libraries, eval=TRUE}
suppressWarnings(suppressMessages(library(stringr)))
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(RCurl)))
suppressWarnings(suppressMessages(library(tidyr)))
```

####Importing the data from Glassdoor
```{r importingdata, eval=TRUE}
joblist <- read.csv("https://raw.githubusercontent.com/delagroove/dataScience/master/jobOffers_final_with_dupes_reduced.csv", stringsAsFactors = FALSE)
```



####Eliminating Duplicate Entries
Since some companies post the same job on different days, the posted date column is elimintated prior to testing for duplicate entries.
```{r eliminate_duplicates, eval=TRUE}
joblisttidy <- joblist[,-5]
joblisttidy <- subset(joblisttidy, duplicated(joblisttidy)==FALSE)
```

####Eliminating symbols in front of location
```{r location, eval = TRUE}
joblisttidy$location <- unlist(str_extract_all(joblisttidy$location, "[[:upper:]]{1}[[:lower:]]{2,}, [[:alpha:]]{2}"))
head(joblisttidy)
```

####Identifying Job Descriptions that look for specific Computer Skills
```{r computerskills, eval=TRUE}
compskills <- joblisttidy %>%
  mutate(python = grepl("python", description, ignore.case=TRUE)) %>%
  mutate(perl = grepl("perl", description, ignore.case=TRUE)) %>%
  mutate(Cplusplus = grepl("C++", description, fixed=TRUE)) %>%
  mutate(SQL = grepl("SQL", description)) %>%
  mutate(java = grepl("java", description, ignore.case=TRUE)) %>%
  mutate(R = grepl(" R,", description)) %>%
  mutate(hadoop = grepl("hadoop", description, ignore.case=TRUE)) %>%
  mutate(spark = grepl("spark", description, ignore.case=TRUE)) %>%
  select(job_title, company, python, perl, Cplusplus, SQL, java, R, hadoop, spark)
summary(compskills)
```

####Identifying analytical skills
```{r analytical-skills, eval=TRUE}
skills <- joblisttidy %>%
  mutate(machinelearning = grepl("machine learning", description, ignore.case=TRUE)) %>%
  mutate(statisticalmodeling = grepl("statistical model", description, ignore.case=TRUE)) %>%
  mutate(techwriting = grepl("technical writing", description, ignore.case=TRUE)) %>%
  select(job_title, company, machinelearning, statisticalmodeling, techwriting)
summary(skills)  
```

####Identifying soft skills
```{r soft-skills, eval=TRUE}
softskills <- joblisttidy %>%
  mutate(collaborative = grepl("collaborat", description, ignore.case=TRUE)) %>%
  mutate(organized = grepl("organized", description, ignore.case=TRUE)) %>%
  mutate(selfstarter = grepl("self starter", description, ignore.case=TRUE)) %>%
  mutate(attndetail = grepl("attention to detail", description, ignore.case=TRUE)) %>%
  mutate(communication = grepl("communicat", description, ignore.case=TRUE)) %>%
  mutate(creative = grepl("creativ", description, ignore.case=TRUE)) %>%
  select(job_title, company, collaborative, organized, selfstarter, attndetail, communication, creative)
summary(softskills)  
```
