---
title: "DATA_607_Project3- Data Science Skills"
authors: "Jill Anderson, Alvaro Bueno Castillo, Nathan Cooper, Silverio Vasquez, Sarah Wigodsky"
date: "October 16, 2017"
output: html_document
---

##What are the most valued data science skills?
To answer this question, we mined the job board glassdoor.com to identify which skills were requested the most frequently.

####Importing Libraries
```{r libraries, eval=TRUE}
suppressWarnings(suppressMessages(library(stringr)))
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(RCurl)))
suppressWarnings(suppressMessages(library(tidyr)))
```

####Importing National Average Data Science Salary By Company
```{r salarydata, eval=TRUE}
salary <- read.csv("https://raw.githubusercontent.com/sjv1030/group_project/master/company_salary.csv", stringsAsFactors = FALSE)
```

####Tidying the Salary Data - Removing dollar sign, commas and making the salary be stored as a number
```{r tidysalary, eval=TRUE}
salarytidy <- salary[,-1]
salarytidy <- salarytidy %>%
  separate(salary, c("salary", "pay_period"), sep="per ")
salarytidy$salary <- gsub(",", "", salarytidy$salary)
salarytidy$salary <- unlist(str_extract_all(salarytidy$salary, "[[:digit:]]{3,}|[[:digit:]]{1,}.[[:digit:]]{1,}"))
salarytidy$salary <- as.numeric(salarytidy$salary)
head(salarytidy)
```

####Converting Daily, Monthly, and Hourly Salaries into a yearly salary assuming a 40 hour work week for 12 months
```{r yearly-salary, eval=TRUE}

#pay_period still says month, day or week even though it is now yearly so that it is clear which salaries were changed so that they are all yearly
for (i in 1:length(salarytidy$salary)){
  if (salarytidy$pay_period[i] == "month") {salarytidy$salary[i] <- salarytidy$salary[i]*12}
  if (salarytidy$pay_period[i] == "week") {salarytidy$salary[i] <- salarytidy$salary[i]*52}
  if (salarytidy$pay_period[i] == "hour") {salarytidy$salary[i] <- salarytidy$salary[i]*8*5*52}
  if (salarytidy$pay_period[i] == "day") {salarytidy$salary[i] <- salarytidy$salary[i]*5*52}
}  
```


####Importing the data from Glassdoor for Boston
```{r importingdataMA, eval=TRUE}
joblistma <- read.csv("https://raw.githubusercontent.com/delagroove/dataScience/master/jobOffers_final_with_dupes_reduced.csv", stringsAsFactors = FALSE)
```

####Importing data from Glassdoor for NYC
```{r importingdataNY, eval=TRUE}
joblistny <- read.csv("https://raw.githubusercontent.com/delagroove/dataScience/master/jobOffers_ny.csv", stringsAsFactors=FALSE)
```


####Importing data from Glassdoor for SF
```{r importingdataSF, eval=TRUE}
joblistsf <- read.csv("https://raw.githubusercontent.com/delagroove/dataScience/master/jobOffers_SF.csv", stringsAsFactors=FALSE)
joblistsf <- joblistsf[,-1]
```

####Combining data tables from Boston, NY and SF
```{r combining-data, eval=TRUE}
joblisttidy <- rbind(joblistma, joblistny, joblistsf)
```

####Eliminating Duplicate Entries
Since some companies post the same job on different days, the posted date column is elimintated prior to testing for duplicate entries.
```{r eliminate_duplicates, eval=TRUE}
joblisttidy <- joblisttidy[,-5]
joblisttidy <- subset(joblisttidy, duplicated(joblisttidy)==FALSE)
```

####Eliminating symbols in front of location
```{r location, eval = TRUE}
joblisttidy$location <- unlist(str_extract_all(joblisttidy$location, "[[:upper:]]{1}[[:lower:]]{2,}, [[:alpha:]]{2}|[[:upper:]]{1}[[:lower:]]{2,} [[:alpha:]]{2,}, [[:alpha:]]{2}"))
```

####Identifying Job Descriptions that look for specific Computer Skills
```{r computerskills, eval=TRUE}
compskills <- joblisttidy %>%
  mutate(python = grepl("python", description, ignore.case=TRUE)) %>%
  mutate(perl = grepl("perl", description, ignore.case=TRUE)) %>%
  mutate(Cplusplus = grepl("C++", description, fixed=TRUE)) %>%
  mutate(SQL = grepl("SQL", description)) %>%
  mutate(java = grepl("java\\b", description, ignore.case=TRUE)) %>%
  mutate(javascript = grepl("javascript", description, ignore.case=TRUE)) %>%
  mutate(R = grepl(" R,", description)) %>%
  mutate(hadoop = grepl("hadoop", description, ignore.case=TRUE)) %>%
  mutate(spark = grepl("spark", description, ignore.case=TRUE)) %>%
  mutate(scala = grepl("scala", description, ignore.case=TRUE)) %>%
  select(job_title, company, python, perl, Cplusplus, SQL, java, javascript, R, hadoop, spark, scala)
summary(compskills)
```

####Identifying analytical skills
```{r analytical-skills, eval=TRUE}
skills <- joblisttidy %>%
  mutate(machinelearning = grepl("machine learning", description, ignore.case=TRUE)) %>%
  mutate(statisticalmodeling = grepl("statistical model", description, ignore.case=TRUE)) %>%
  mutate(techwriting = grepl("technical writing", description, ignore.case=TRUE)) %>%
  mutate(plateau = grepl("plateau", description, ignore.case=TRUE)) %>%
  mutate(d3 = grepl("D3", description)) %>%
  select(job_title, company, machinelearning, statisticalmodeling, techwriting, plateau, d3)
summary(skills)  
```

####Identifying soft skills
```{r soft-skills, eval=TRUE}
softskills <- joblisttidy %>%
  mutate(collaborative = grepl("collaborat", description, ignore.case=TRUE)) %>%
  mutate(organized = grepl("organized", description, ignore.case=TRUE)) %>%
  mutate(selfstarter = grepl("self starter", description, ignore.case=TRUE)) %>%
  mutate(attndetail = grepl("attention to detail", description, ignore.case=TRUE)) %>%
  mutate(communication = grepl("communicat", description, ignore.case=TRUE)) %>%
  mutate(creative = grepl("creativ", description, ignore.case=TRUE)) %>%
  mutate(visualization = grepl("visualization", description, ignore.case=TRUE)) %>%
  select(job_title, company, collaborative, organized, selfstarter, attndetail, communication, creative, visualization)
summary(softskills)  
```
