---
title: "DATA 607 HW 9"
author: "Sarah Wigodsky"
date: "October 27, 2017"
output: html_document
---
##Introduction
####I chose to search the New York Times Article Search API for mentions of the word secession between December 1, 1860 and December 7, 1860.

####Load Libraries
```{r load-libraries, eval=TRUE}
suppressWarnings(suppressMessages(library(httr)))
suppressWarnings(suppressMessages(library(RCurl)))
suppressWarnings(suppressMessages(library(tidyr)))
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(rtimes)))
suppressWarnings(suppressMessages(library(readr)))
```

####Reading in the API Key.  It is stored in a file on my computer so that it remains secure.
```{r APIkey, eval=TRUE}
NYTIMES_AS_KEY = read_file("C:/Users/Swigo/OneDrive/Documents/GitHub/DATA-607/nytimesapi.txt")
```

####Searching the New York Times Article Search API
#####Searching for the word secession between the dates of 12/1/1860 and 12/7/1860
```{r article-search, eval=TRUE}
secession_search <- as_search(q="secession", begin_date = "18601201", end_date = "18601207", key=NYTIMES_AS_KEY)
```

####Selecting the Publication Date, Type of Article, Headline, Web URL
```{r selecting-columns, eval=TRUE}
nytimes_secession <- select(secession_search$data, 7, 9, 13, 1)
```

####Converting the Tibble into a Dataframe
```{r dataframe, eval=TRUE}
nytimesdf <- data.frame(nytimes_secession)
nytimesdf
```